---
title: "The importance of measurement uncertainty in ecological management"
author:
  - name: "Milad Memarzadeh"
    affiliation: a
  - name: "Carl Boettiger"
    affiliation: a, 1
address:
  - code: a
    address: "Dept of Environmental Science, Policy, and Management, University of California Berkeley, Berkeley CA 94720-3114, USA"
  

abstract: |
  We settle a long standing paradox in ecological management regarding the influence of measurement uncertainty.  Previous work has frequently suggested that the optimal management of fisheries should respond to increased measurement uncertainty by increasing harvest effort, contrary to intuition and the precautionary principle. Through the use of cutting-edge point-based partially observed Markov decision process (POMDP) methods, we demonstrate how this paradox arises from flaws in the standard management theory, which contributes to over-exploitation of fisheries and increased probability of economic and ecological collapse.  In contrast, we find POMDP-based management avoids such over-exploitation while also generating higher economic value through adaptive learning. 


keywords:
  - Measurement uncertainty
  - POMDP
  - Decision theory
  - Fisheries
  - Conservation

lineno: true

bibliography: citations.bib
csl: the-american-naturalist.csl
output: rticles::elsevier_article
layout: 3p # review # review = doublespace, 3p = singlespace, 5p = two-column

---


```{r libraries, include=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning=FALSE, fig.width=7, fig.height=7)
library(tidyverse)
library(gridExtra)
```

```{r optional_plot_fonts, message=FALSE}
library(ggplot2)
library(Cairo)
library(extrafont)
library(hrbrthemes)

extrafont::loadfonts()
knitr::opts_chunk$set( dev="cairo_pdf")
hrbrthemes::import_roboto_condensed() 
ggplot2::theme_set(hrbrthemes::theme_ipsum_rc())
```

```{r plot_colors}
#palette <- ggthemes::ptol_pal()(6)
palette <- c("#D9661F", "#3B7EA1",  "#6C3302", "#FDB515", "#00B0DA",  "#CFDD45") # berkeley colors

colors <- set_names(c(palette[c(1:4,2,5)], "grey", "black"), 
                    c("TAC", "POMDP", "MSY", "CE", 
                      "POMDP: low prior",
                      "POMDP: medium prior",
                      "biomass",
                      "catch"))

```


```{r load_data, message=FALSE, warning=FALSE}
log_dir <- "../data" # solution files here

sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/sims.csv"), col_types = "inicnnn")
det_sims <- read_csv(file.path(log_dir, "observe_harvest_recruit/det_sims.csv"))
all_policies <- read_csv(file.path(log_dir, "observe_harvest_recruit/all_policies.csv"))
ram_ex <- read_csv(file.path(log_dir, "ram_ex/ram_ex.csv"))

```


Imperfect information is ubiquitous in ecological management and conservation decision making. While the pressing concerns of global change have put the spotlight on *forecasting*, i.e. the uncertainty of the future [e.g. @Petchy2015], management decisions must also contend with uncertainty of the *present*: How many fish are in the sea today? What regions harbor the most fragile biodiversity? How have invasive species transformed native communities? A recognition of the importance of uncertainty as well as sophisticated methods and tools to do so has become increasingly widespread in the context of estimating ecological models from available data, thanks in large part to the rise of approaches such as Hierachical Bayesian Modeling, along with convenient software tools for implementing these methods.  In sharp contrast to that shift, the treatment of uncertainty in decisions -- how the uncertainty in models should be reflected in scientific recommendations, remains far less developed, more qualitative, and underserved by available tools.  In this paper we illustrate how uncertainty in measurement has a dramatic negative impact on outcomes under a variety of these exsiting approaches, which can be avoided through the use of more formal approaches that have since been developed in other fields that can likewise be implemented in software tools.  This insight both resolves a long-standing paradox in fisheries management and has important implications for how ecosystem management deals with uncertainty more generally.

Approaches for decision making under uncertainty in ecological systems can be divided into two camps: approaches based in "optimal control" and usually favored by natural resource economists, and approaches based in heuristic methods such as scenario planning, resilience thinking, and precautionary rules of thumb, more commonly found in both ecological literature and actual practice [@Fischer2009; @Polasky2011].  While researchers have for some time recognized the need to unify the transparent and quantitative algorithmic approach of optimal control with the greater complexity and uncertainty of real ecosystems that is acknolwedged by heuristic methods, computational barriers to doing so have hither-to stymied this progress.  Here, we illustrate how the limitations of these approaches has been manifested in the example of fisheries management, and present a new approach that can combine the reality of measurement uncertainty and the rigour of optimization to resolve a long-standing paradox and suggest a more robust approach to management.   

Fisheries conservation and management has long been both a crucible and proving ground for the theory of ecological management more generally, including topics such as adaptive management [@Walters1978], ecosystem-based management [@Levin2008] and resilience thinking [@Holling1973; @May1977], while also giving rise to the sub-discipline of resource economics [@Beverton1957; @Schaefer1954; @Gordon1954], thanks to its global relevance, long history, and readily available data.  Consequently, understanding the impact of measurement uncertainty in this rich and well-studied context will have implications for both resource management and conservation efforts in other domains, many of which draw on concepts established or tested in fisheries ecosystems.  In particular, the approach here illustrates the limitations of both over-simple optimal control approaches and rule-of-thumb precautionary approaches, and offers instead a more general strategy for dealing with measurement uncertainty adaptively.  


We compare the results of managing simulated fish stocks under the existing optimal control approach, a heuristic approach that more closely resembles today's management practices, and a newly proposed approach that provides a more rigorous treatment of uncertainty.

# Model

We consider the management problem of setting catch quotas for a marine fishery in the face of imperfect information about the current stock size and uncertainty about future recruitment.  To facilitate tractability and intepretation, we will focus on the well-studied Gordon-Schaefer model [@Schaefer1954; @Gordon1954] of logistic population growth (stock recruitment),  

$x_{t+1} = x_t r  \left(1 - \frac{x_t}{K}\right) - h_t \varepsilon_t$

where $x_t$ is the current stock size, $h_t$ the harvest chosen that year, with parameters $r$ giving the individual growth rate, and $K$ the carrying capacity and $\varepsilon$ representing stochastic recruitment in a variable environment.  We will assume for simplicity $\varepsilon \sim \mathcal{N}(1, \sigma_g)$.  

This and similar models are widely used in large scale analyses across diverse stocks (e.g. [@Costello2016; @Britten2017]), and form the basis for much of bioeconomic theory [@Clark1990]. As such, it will be easier to compare our results against intuition and classic theory, and avoid the possibility of differences arsing only because of some particular subtle assumption hidden in a more complex model.

However, it is important to remember that both existing methods we discuss and the POMDP approach proposed here can or are already applied to more complex models as well, including those with age or stage structure.  Likewise, while we consider the model and it's parameters as "given," it should be understood to mean that this in practice comes as the result of some model choice and statistical estimation from historical data, frequently accounting for uncertainty in both the process and the measurements [e.g. @].  As we shall see here, having included uncertainty in the model *estimation* in no way excuses us from having to also deal once again with that uncertainty in the decision process.  

# The Decision Problem

We seek to determine the sequence of actions $h_t$ for $t \in [1, ... \infty]$ that maximize the net present value (discounted sum of all future profits) of the fishery.  We will denote the discount rate $\gamma$. For simplicity we will once again follow classic theory and assume a fixed price for fish, (equivalently, measuring our value in units of discounted fish rather than discounted dollars), though again this assumption is easy to relax. Each year, the manager also obtains an estimate $y_t$ of the true $x_t$ stock size,

$$y_t = \xi_t x_t$$

Given an estimated recruitment model along with this measurement of the stock size, the manager must set a harvest quota $h_t$ for that year in order to maximize their expected long term (net) present value:

$$\max_{ \lbrace h_t \rbrace} \mathbb{E} \left \{ \sum_{t=0}^{\infty} \gamma^t \cdot h_t \right \}$$

A few subtlies arise in this seemlingly simple problem statement that we must address.  First, given our discrete-time formulation of the recruitment process, it is necessary to decide if the measurement $y_t$ happens before or after harvest $h_t$: that is do we: measure, recruit, harvest, or measure, harvest, recruit?  Following convention in the optimal control literature (e.g. [@Reed1979; @Clark1990; @Sethi2005]), we will assume the latter; measurement occurs immediately before harvest, followed by recruitment. 

A second subtly is in the nature of the decision problem itself: the optimization problem must always consider the set ${h_t}$ all future actions simultaneously, not merely the next action.  This is a game of chess, where it can be optimal to make costly short term actions (such as closing the fishery) if they pay off in the long run. This differs from the problem solved in theory as well as the practice of maximum sustainable yield, which seeks to identify only a single harvest quota $H_{MSY}$ applied for all time, rather than a *policy* of varying quotas $h_t$ depending on the stock assessment $y_t$.  Understanding this distinction is key both to understanding the differences between current theory and current best practice in fisheries management.


<!-- MDP, POMDP -->

# Current theory and practice

@Gordon1954 and @Schaefer1954 independently showed in the same year that the maximum sustainable yield for the model bearing their names is achieved by reducing the stock to the  population to the size at which it obtains the maximum growth rate which is known as Biomass at Maximum Sustainable Yield, $B_{MSY}$. For the Gordon-Schaefer model (and many others), this is achieved at $B_{MSY} = K/2$.  They observed that fishing at a *constant yield* (that is, an individual fish mortality, $F$, such that harvest $H = F \cdot B$) of $F_{MSY} := \tfrac{H_{MSY}}{B_{MSY}}$)  will eventually lead to a population that converges to the biomass $B_{MSY}$ and produces the maximum sustainable harvest, $H_{MSY} = r K / 4$ in this model.
From here, theory and practice diverge.  

This *constant yield* (constant mortality) solution thus corresponds to an equlibrium analysis, and does not solve the time-dependent optimization problem above. In particular, maximum sustainable yield will not be optimal whenever the stock is away from $B_{MSY}$.  @Clark1973 shows that the optimal time-dependent strategy is not one of constant yield, but rather of *constant escapement*. The time-depednent optimal stategy is impatient: it seeks to achieve the optimal yield as quickly as possible.  Thus, at any level below $B_{MSY}$, the economically optimal thing to do is to completely shut down the fishery, while above this stock, mortalities greater than $F_{MSY}$ can be optimal to bring the stock back to $B_{MSY}$ more quickly.  (This intution must be adjusted slightly in the case of economic discounting of future profits, but is otherwise quite general, see @Clark1990).  This strategy is known as constant escapement, since a constant sock size $B_{MSY}$ escapes harvest each year.  While both converge to the same long-term biomass and long-term yield, these differences away from equilibrium between the mortality-based approach that dominates fisheries management practice and the escapement-based approach that became the focus of theory will result in different behaviors under uncertainty.




The literature of optimal harvests under uncertainty is one of paradoxes, which can only be resolved by a more comprehensive treatment of measurement uncertainty.  

## Reed's Paradox: $S = D$

The first of these we shall refer to as Reed's Paradox, owing to the mathematical proof provided in @Reed1979 which demonstrates that, under sufficiently general assumptions, the optimal escapement $S$ for a population under stochastic growth, is identical to the optimal escapement $D$ of a deterministic population, $S = D = B_{MSY}$.  This suprising result suggests that in going from a world where a manager has perfect knowledge of absolutely everything into a scenario where the manager faces considerable uncertainty about the future state of the world, no additional precaution is needed.  This also provides the manager with a remarkably convenient mechanism for determining the optimal harvest policy: rather than rely on computationally intensive Stochastic Dynamic Programming to determine the optimal policy, a simple derivative will suffice, which is all that is needed to solve the maximum of the deterministic problem.  

## Clark's Paradox: $M < D$

@Clark1986 was among the first attempts to resolve Reed's Paradox. @Clark1986 (quite correctly, as we will see), identified the crux of Reed's Paradox as the absence of measurement uncertainty:

> An important tacit assumption in Reed’s analysis, as in the other works referred
to above, is that the recruitment level X is known accurately prior to the harvest decision, [...] In the case of fishery resources, the stock level X is almost never known very accurately, owing to the difficulty of observing fish in their natural environment.

Unfortunately, @Clark1986 was unable to solve the resulting problem exactly, but had to adopt an almost equally troublesome assumption:

> For reasons of tractability, we shall adopt the simplifying assumption that the escapement level S, is known exactly at the end of that period. (The mathematical difficulty
of the problem increases markedly if this assumption is relaxed.) 

By leveraging algorithms (and also computational hardware) developed over two decades later are we able to crack this mathematical difficulty Clark refers to directly without this "simplyfying assumption."  Unforunately for @Clark1986, this "simplifying assumption" they made instead only deepened the paradox, leading them to find policies that become *less* cautious as uncertainty increases:

> [Our] results appear to contradict the conventional wisdom of renewable resource management, under which high uncertainty would call for increased caution in the setting of quotas.

Using a different but still flawed assumption nearly two decades later, @Sethi2005 largely confirm Clark's Paradox, which they likewise observed with some concern:

> It may seem counter-intuitive that a measurement error causes lower expected escapements below the deterministic fishery closure threshold.

While other attempts to introduce measurement error have been considered, each relies on various approximations that similarly alter the problem, such as assuming the solution must be of 'constant-escapement' type [@Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].  Here we will illustrate the use of powerful modern algorithms that will permit a more direct solution to the case of measurement uncertainty.  






First we will introduce the more heuristic alternative to addressing uncertainty, which more accurately reflects both intuition and actual management.  

## Uncertainty in Practice: TAC

Fortunately, management practice has so far ignored this paradoxical suggestion that uncertainty should be met with greater rather than reduced fishing quotas. Maximum Sustainable Yield (MSY) remains the basis of international law (including the UN,  IWC, IATTC ICCAT, ICNAF; @Mace2001) and a familiar standard of management in terrestrial ecosystems as well as aquatic [@Clark1990].  Critics have for some time observed the limitations of harvesting at MSY in face of uncertainty in stock sizes and population dynamics [@Larkin1977; @Botsford1997].  Many US fisheries reflect this uncertainty through a series of adjustments that effectively reduce the target fishing mortality level to reflect this uncertainty.  Typically, a stock assessment model provides a best estimate of $B_{MSY}$ and corresponding mortality $F_{MSY}$ is used to define the stock Overfishing Limit, (OFL). Based on this, a somewhat lower level is set as the Allowable Biological Catch (ABC), reflecting uncertainty in the stock assessment. To reflect possible uncertainty between reported and actual catch, the ABC is reduced somewhat further to define the Total Allowable Catch, (TAC), which forms the basic unit of management for many such fisheries.  To reflect this process, our analysis will also consider policies in which the harvest quota is set at 80% of the level expected under and MSY policy: $H_{TAC}(t) = 0.8 \cdot F_{MSY} \cdot B_t$, for a biomass estimated at $B(t)$ in year $t$. To distinguish this approach from MSY, we will refer to this approach as a TAC policy.  This more closely represents a heuristic [e.g. @Hilborn2010] or resilience-based approach [sensu @Fischer2009] than an optimization-based policy.  Importantly, this approach shares the fundamentally stationary assumptions of an MSY policy by defining a constant mortality rather than a dynamic policy.  Thus, despite being more cautious overall and generating lower economic yield than a constant escapement policy, TAC policies continue to harvest at non-zero rates even if a stock falls below $B_{MSY}$, while the constant escapement policy does not.  



## POMDPs: An optimal treatment of measurement uncertainty

As discussed above, absent the short-cut provided by Reed's theorems -- solving for the optimal management policy in a stochastically fluctuating population -- involves a Markov Decision Process (MDP), commonly referred to in ecological literature by its solution method, Stochastic Dynamic Programming (SDP). Once measurement uncertainty is introduced, the problem becomes a Partially Observed Markov Decision Process (POMDP, not be confused with a partially observed or hidden Markov model, HMM, in which no decision process is involved).  The POMDP problem is posed identically to that of the MDP problem, with the addition of an observation process.  The POMDP problem for fisheries question considered here can be summarized as follows:

- Transition function (state equation): $T(x_t, x_{t+1}, a_t)$: the probability that a system is in state $x_{t+1}$ at time $t+1$ given that it began in state $x_t$ at time $t$ and the manager took action $a_t$.  In our context, this relationship is given by the Graham-Schaefer stock recruitment function $f$ with normally distributed growth uncertainty $\mathcal{N_T}(f(x_t,a_t), \sigma_g)$, truncated at zero to exclude negative population sizes.

- Observation function: $O(x_t,y_t,a_t)$ the probability of observing state $y$ given a system in state $x$.  In principle, the action chosen can influence the precision of the observation. In our case, we simply assume normally distributed errors around the true state, $y \sim \mathcal{N_T}(x, \sigma_m)$, truncated at zero to exclude negative population sizes.  

- Utility function: $U(x_t,a_t)$, the utility recieved at time $t$ for taking action $a_t$, given that the system is in state $x_t$.  For simplicity of analysis, we will simply set the utility to be equal to the harvested stock: $U(x_t, a_t) = \min(x_t - a_t, 0)$, indicating that realized harvest cannot be negative.  This choice ensures that in the case of no uncertainty ($\sigma_g = \sigma_m = 0$), the optimal solution matches that expected under a simple MSY calculation. More realistic utility functions may include diminishing returns with increasing harvest (supply and demand effects), and the cost of fishing, both of which act to suppress large harvests. By focusing on a simple utility we can be sure that our comparison to MSY is driven by the treatment of uncertainty rather than merely differing economic assumptions.  

The optimization problem is to select the action $a_t$ that will maximize the net present utility over all time. Future utility may be discounted by a factor $\gamma$, so that a value  $V$ in $t$ year is valued at $\gamma^t V$ today.  


Numerically, each of these functions are defined over a discrete set of possible states, observations, and actions, and can thus be represented as a collection of matrices or tensors.  As in the case of MDP [@Marescot2013], the POMDP problem can be solved by existing and well-studied algorithms. While not unknown to the conservation literature [@Williams2011], algorithms for POMDP have historically scaled quite poorly, and their application has been restricted to contexts with only a handful of possible states and actions [e.g. @Chades2008; @Chades2011; @Fackler2014; @Fackler2014b].  POMDP algorithms remain a very active area in artificial intellegence, and by adapting recent algorithmic developments [@Kurniawati2008] from that field we are able to find solutions for the considerably more complex problems such as the case of fisheries managment, which requires on the order of 100 states and actions to provide sufficient numerical resolution. Readers interested in the technical details of these algorithms should consult the extensive literature in in POMDP methods. For those interested in exploring or applying these methods, we have developed a preliminary R package [@sarsop-pkg], which can be used to replicate any of the results presented here and explore other variations.  

# Results 

<!--
TAC, MDP, POMDP (sigma_g, sigma_m)
-->



Figure 1 shows average fish biomass across 100 replicate simulations under three different management strategies: constant escapement (CE), total allowable catch (TAC; equal to 80% MSY), and partially observed Markov decision process (POMDP) management.  Each successive panel shows a subsequently higher level of measurement error, from $\sigma_m = 0$, $\sigma_m = 0.1$, and $\sigma_m = 0.15$, as indicated.  Standard deviation from the mean across replicate simulations is shown as faint colored bands, indicating significant variation due to stochasticity between individual replicates.  In each panel shown in Figure 1 stochastic recruitment (environmental noise) is set to a moderate $\sigma_g = 0.15$.  In the absence of either environmental noise or measurement error (Supplemental material, Figure S1) POMDP, CE, and MSY would converge to stock at the $B_{MSY}$, while TAC would maintain the stock at a slightly higher level.  The first panel of Figure 1 shows that the the introduction of stochastic growth has a significant negative impact on the TAC strategy (This impact is even more severe for MSY, which always harvests strictly more than TAC by definition, as, shown in Supplemental material Figure S2), which results in an average biomass significantly lower than $B_{MSY}$.  Without mesasurement error, the CE and POMDP strategies are nearly identical with both approximately maintaining the stock at $B_{MSY}$ despite the significant environmental stochasticity.  As measurement error increases in the subsequent panels, CE and TAC strategies perform increasingly poorly, while the POMDP continues to maintain the average stock close to $B_{MSY}$.  Notably, CE is significantly more impacted by measurement uncertainty than TAC, with CE averaging even lower biomass than TAC under moderate measurement uncertainty.  

Figure 1 confirms that the precautionary approach represented by the TAC does indeed prove more robust to the problem of measurement error than the optimization solution represented by CE.  In general, we expect that the rigid assumptions required of optimization will lead to worse outcomes when those assumptions are not met than we see under the corresponding heuristic approach.  Yet it is important to bear in mind that the opposite pattern was observed in the case of environmental stochasticity, in which it was the heuristic TAC stategy rather than the CE strategy that lead to the highest overfishing rate.  In contrast, the POMDP approach successfully handles both stochasticity and observation error, maintaining the stock near $B_{MSY}$ levels that produce the highest yield.  

```{r fig.width=7, fig.height=3.5, fig.cap="Average fish biomass under different management strategies under increasing levels of measurement uncertainty. Each plot the mean stock size over time across 100 replicate simulations under each policy: constant escapement (CE), Total allowable catch (TAC = 80% MSY), and the proposed partially observed Markov decision process (POMDP) method. Measurement error increases as a normal distribution with standard deviation 0, 0.1, or 0.15, as indicated at the top of the panel.  Environmental stochasticity is fixed a standard deviation of 0.15 in each panel.  Carrying capacity K normalized to 1, r = 0.75. Additional environmental noise levels and comparison to MSY rather than TAC can be found in the supplementary material."}
sims %>%
  select(-value) %>%
  filter(method != "msy", sigma_g == "0.15") %>%
  mutate(method = fct_recode(method, 
                             "CE" = "det",
                             "TAC" = "pgy",
                             "POMDP" = "pomdp")) %>%
  group_by(time, method, sigma_m) %>%
  summarise(mean = mean(state), sd = sd(state)) %>% 
  mutate(sigma_m = as.character(sigma_m)) ->
fig1_df


fig1_df %>%
  ggplot(aes(time, mean, col=method, fill=method)) + 
  geom_line(lwd=1) + 
  geom_ribbon(aes(ymax = mean + sd, ymin = mean-sd), col = NA, alpha = 0.1) +
  facet_wrap(~ sigma_m, labeller = label_bquote(sigma[m] == .(sigma_m))) + 
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  coord_cartesian(ylim = c(0, 1)) + 
  theme(legend.position = "bottom") + 
  ylab("Mean biomass")
```

This pattern in management success in ecological terms (the relatively recovery and maintenance of fish biomass) is also bourne out in terms of economic performance.  Figure 2 shows the mean net present value (averaging across replicates and discounting future profits by the discount factor $\gamma$) for these same simulations at increasing levels of measurement uncertainty.  As before, environmental stochasticity is set at $\sigma_g = 0.15$.  In the absense of measurement uncertainty, constant escapement (CE) is optimal, though POMDP performs nearly as well, while TAC underperforms.  As measurement uncertainty increases, the economic value of CE declines, remaining better than TAC at modest measurement error $\sigma_m = 0.1$ but worse at higher uncertainty.  Though the risk of stock collapse does increase with increasing measurement error under TAC (as seen by the mean declines in Figure 1), these have little impact on the economic value due to the discount rate.  A smaller discount rate would penalize unlikely but not improbable collapses more, since a significant amount of time is required to realize those rare events.  Supplemental Figure S3 summarizes these economic trends across different noise values $\sigma_g$ and includes comparison to a simple MSY policy.


```{r fig.width=7, fig.height=3.5, fig.cap=""}
sims %>%
  select(time, value, rep, method, sigma_m, sigma_g) %>%
  filter(sigma_g == "0.15", method != "msy") %>%
  group_by(rep, method, sigma_m, sigma_g) %>%
  summarise(npv = sum(value)) %>%  
  group_by(method, sigma_m, sigma_g) %>%
  summarise(net_value = mean(npv), se = sd(npv) / mean(npv)) %>%
  ungroup() %>%
  mutate(method = fct_recode(method,
                             "CE" = "det",
                             "TAC" = "pgy",
                             "POMDP" = "pomdp")) %>%
  
  
  ggplot(aes(method, net_value, ymin=net_value-se, ymax=net_value+se, fill=method)) + 
  geom_bar(position=position_dodge(), stat="identity") + 
  geom_errorbar(size=.3,width=.2,
                position=position_dodge(.9)) +
  facet_wrap(~sigma_m, 
             labeller = label_bquote(sigma[m] == .(sigma_m))) +
  theme(legend.position = "none") + 
  scale_color_manual(values = colors) +
  scale_fill_manual(values = colors) +
  ylab("expected net present value")
```



<!-- Show results of POMDP when over-estimating! -->


To better understand the differences in performance of these strategies and resolve the paradox of uncertainty, we must take a closer look at how the specific action recommended by each policy compares given the same observation.  Figure 3 shows the action taken by each strategy in response to a measurement of the stock size.  The first panel shows the policy in terms of harvest $H$, that is, the catch quota that would be set by each strategy.  For reference, the second panel shows the same policies but in terms of expected escapement, $S = Y - H$ for observed stock $Y$, which is more common in the control literature [e.g. @Reed1979; @Clark1987; @Sethi2005] than in fisheries literature.  The first panel clearly illustrates the contrast between the constant escapement (CE) strategy and the precautionary Total Allowable Catch (TAC) strategy: CE sets harvest strictly to zero for stocks estimated at biomass below $B_{MSY}$, while TAC permits a modest harvest of even very small stocks.  Viewing the same policies in terms of escapement, second panel, makes it clear that only CE strategy corresponds to keeping escapement constant.  This is important, as theoretical work ever since the proofs of @Reed1979 often attempts to simplify the calculations by assuming the optimal policy must be a constant-escapement type strategy [e.g. @Ludwig1981; @Roughgarden1996; @Engen1997;  @Moxnes2003].  

While the catch quota under both TAC and CE stategies can be completely determined given the most recent observation of the stock size by using the policy curves shown in Figure 2, this is not the case for the POMDP approach. This fundamental difference is key to understanding the difference in performance and resolving the paradox of uncertainty.  The POMDP policy cannot be specified by the most recent observation alone. Instead, the POMDP policy depends on all prior observations, not just the most recent.  The reason for this complexity comes from the Markov property.  Observations of the state in the perfectly observed system satisfy the Markov property: once we have measured the current biomass exactly, we cannot get any better estimate of the current stock size by studying older measurements. When measurements are uncertain this is no longer the case: intuitively, by comparing the most recent measurement to previous observations we may be able to infer when any given measurement is unusually high or unusually low.  POMDP formalizes this intutition by capturing the information from all previous observations into a *prior belief*.  This prior belief is updated after every subsequent action and observation in accordance with Bayes Law.  The mechanics of this process are well documented in the extensive literature on POMDPs [e.g. ], but for our purposes 

The other approaches exploit this property, and the simplifying assumptions introduced in previous work attempting to address measurement uncertainty [@Clark1987; @Sethi2005] serve to restore or assert this.  The POMDP 


It is important to bear in mind several issues when intepreting these figures. First, the horizontal axis reflects the measured stock size, which may differ from the (unobservable) true stock size when measurement uncertainty is present.  

Consequently, the realized escapement may differ from the target level.  



POMDP policies depend on the prior belief.  

```{r fig.width=7, fig.height=7, fig.cap=""}
all_policies %>% 
  filter(policy %in% c("low", "medium", "det", "pgy")) %>%
  mutate(policy = fct_recode(policy,
                             "CE" = "det",
                             "TAC" = "pgy",
                             "POMDP: low prior" = "low",
                             "POMDP: medium prior" = "medium"
                             )) -> policy_df
harvest <- policy_df %>%
  ggplot(aes(state, harvest, col=policy)) + 
  geom_line(lwd=1)  +
  scale_color_manual(values = colors) +
  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) + 
  ggtitle("Harvest view") + 
  theme(legend.position="none") + 
  xlab("Estimated stock biomass")

escapement <- policy_df %>%
  ggplot(aes(state, escapement, col=policy)) + 
  geom_line(lwd=1)  +
  scale_color_manual(values = colors) +
  coord_cartesian(xlim = c(0,1), ylim = c(0, .8)) +
  xlab("Estimated stock biomass") +
  ggtitle("Escapement view") + 
  theme(legend.position = "bottom")

gridExtra::grid.arrange(harvest, escapement, 
                        layout_matrix = rbind(c(1,1), c(2,2)))

```




```{r message=FALSE, warning=FALSE, fig.width=7, fig.height = 3.5, fig.cap="Comparisons of harvest level that would be recommended by policies considered here relative to historical harvest levels in two commercial fish stocks.  POMDP solution assumes a measurement error of 10%."}
ram_ex %>%
  filter(model %in% c("biomass", "catch", "POMDP", "Reed", "pgy")) %>%
   mutate(model = fct_recode(model,
                             "CE" = "Reed",
                             "TAC" = "pgy"
                             )) %>%
  ggplot(aes(year, stock, col=model)) +
  geom_line(lwd=1) +
  scale_color_manual(values = colors) +
  facet_wrap(~commonname, scales = "free", ncol=1) + theme_bw()  
```


**A POMDP policy reflects all prior observations**.
The essential difference in a POMDP policy compared to either that of constant escapement $S$ or constant mortality $F$ is that nothing is just a constant -- instead, the harvest quota recommended by POMDP will always depend not only on the most recent observation of the stock size, but on all previous stock sizes.  This dependency can be expressed in terms of a *prior belief* -- if all prior observations have led one to expect that the stock should be relatively small, a single observation of a very large stock is met by some skepticism under a POMDP, and the resulting harvest recommended by POMDP can differ substantially from that recommended by constant escapement or constant mortaility, which by definition consider only the most recent observation.  If the observation is largely consistent with prior expectations, a POMDP policy will be more similar to that of the constant escapement policy, though with a natural level of precaution built in to allow for the possibility of error in measurement coupled with a negative shock to growth.  


# Discussion

Is it surprising that POMDP model should do better?  Recall that Reed found stochasticity to make no difference, the deterministic model performs identically.  

## A practical implementation?


The year-to-year variation in quotas necessary under either a constant escapement policy
or a policy in which quotas are set by POMDP can seem impractical from a top-down
management perspective, particularly when compared to the simplicity of fixing a mortality of fishing $F_{MSY}$.
However, the economist's perspective reveals that this may not be as unrealistic as it may first appear.
Fisheries today are not managed at either their maximum ecological or economic potential.  
Taken as a whole, most ocean fisheries yield a net negative economic return
[@FAO2009], with fishing persisting only with the support of government subsidies [@Arnason2012], 
meanwhile some 68% of global fisheries are not in good biological condition [@Costello2016].
While the causes for this are both complex and numerous, we have seen that both the economic
inefficiency and ecological health are hurt by current policies, including
both MSY and it's more precautionary variants, as these all permit some level of harvest
below $B_{MSY}$.  Economists have for some time observed these inherent 
inefficiencies, and argued that the establishment of property rights in fisheries [@Arnason2012].
Approaches such as individual transferable quotas (ITQs) would incentivise
fisheries behavior to converge towards the more dynamic harvests of a constant
escapement policy, as doing so would increase the value of an individual's share.
@Arnason2012 argues that properly implemented, this would lead fisheries to desire
an even more dynamic adjustment of the TAC to maximize the value of their shares,
rather than to oppose the  @Costello2016 refers to this as Rights Based Fisheries 
Mangement, (RBFM), and it is precisely the willingness to shut down harvests for 
over-exploited fisheries that explains the faster recovery they predict under RBFM
relative to MSY. Though the @Costello2016 RBFM projections are based on a simple 
constant escapement policy which ignores uncertainty, the logic is quite general:
given property rights to establish an efficient market, it is not unrealistic to
expect fisheries adopting such dynamic polices. Indeed, the greater adoption of 
RBFM should spur the adoption of POMDP approaches as rights-holders seek strategies
that maximize the share values.  This is not to dismiss the immense practical
challenges in such an implementation, but to observe that such dynamic, better-performing
policies can be seen to embrace rather than ignore economic realities.  Notably,
the optimization approach assumed in POMDP can be easily adapted to reflect other
constraints such as the cost of idle fleets if a fishery is closed, or include
the value of ecosystem services from fish not harvested [@Halpern2013].  




<!--Alternate utility functions do not increase
the complexity of the solution.  Note that fixed cost of a fleet doesn't alter the answer at all:
if you have the strategy that optimizes your profits when you have no expenses, you don't do a 
different strategy that gets less profit unless it will also lower costs.  So whether or not
you have to pay a fixed fee each year for boats, even if those boats are idle, is no reason
not to leave those boats idle.  only if leaving them idle is *more expensive* than operating them 
does this make sense, which seems unlikely.  -->

<!-- True whether or not it is implemented in effort-based control or as a TAC-->


<!--
Moreover, regulating fishing effort rather than fishing catch, i.e. opening
a fishery for only certain days of the year rather than limiting dockside catch, is both
economically less efficient: for instance, it increasing costs due to the incentive to use more or larger fishing boats
to maximize catch in that time, and it decreases profits since much of the year's catch
is landed in a short window where it drives down prices, instead of being sold fresh throughout
the year. Because regulating effort is both less direct and more difficult to enforce, the uncertainty
in the expected harvest is also larger than in regulating catch.  
-->

# References 


